---
title: "Tennis HMMs"
output: html_notebook
---

#Federer's Matches in Aus Open 2017

```{r loaddata}
setwd("~/Documents/Winter Project/intentions/data")
library(devtools)
install()
load("/Users/Jeremy/Documents/Winter Project/intentions/data/federer2017.RData")

library(ggplot2)
library(tidyr)
library(dplyr)
library(readr)
```

```{r df}
fed_df <- federer2017

fed_df <- arrange(fed_df, matchid, set, game, point, shot)


fed_df <- fed_df %>%
  mutate(lastshot = ifelse(fed_df$shot == fed_df$final.shot,1,0)) %>%
  mutate(fedhit = ifelse(fed_df$impact.player == "FEDERER",1,0)) %>%
  mutate(isserver = ifelse(fed_df$server == fed_df$impact.player,1,0)) %>%
  mutate(fhand = ifelse(hitpoint == "F",1,0)) 

fed_df <- fed_df %>%
  mutate(opponent = ifelse(fed_df$server == "FEDERER", fed_df$receiver, fed_df$server))

fed_df <- fed_df %>%
  mutate(winner = ifelse(fed_df$ended.in.error == FALSE & fed_df$lastshot == 1,1,0)) %>%
  mutate(retser = ifelse(fed_df$shot == 2,1,0))

for (i in 1:ncol(fed_df)) {
  if(is.character(fed_df[,i]) == TRUE) {
    fed_df[,i] <- as.factor(fed_df[,i])
  }
}
```

Splitting id into sets won, games won, points won

```{r scores}
library(stringi)

ids <- fed_df$id
split.ids <- t(data.frame(stri_split_boundaries(ids, type="character")))
fed_df <- fed_df %>%
  mutate(server.points = split.ids[,1], receiver.points = split.ids[,2], server.games = split.ids[,3], receiver.games = split.ids[,4], server.sets = split.ids[,5], receiver.sets = split.ids[,6])

#Now mutate this to Federer score and opponent score
fed_df <- fed_df %>% rowwise %>%
  mutate(fed.points = ifelse(server == "FEDERER", as.numeric(server.points), as.numeric(receiver.points))) %>%
  mutate(fed.games = ifelse(server == "FEDERER", as.numeric(server.games), as.numeric(receiver.games))) %>%
  mutate(fed.sets = ifelse(server == "FEDERER", as.numeric(server.sets), as.numeric(receiver.sets))) %>%
  mutate(opp.points = ifelse(server == "FEDERER", as.numeric(receiver.points), as.numeric(server.points))) %>%
  mutate(opp.games = ifelse(server == "FEDERER", as.numeric(receiver.games), as.numeric(server.games))) %>%
  mutate(opp.sets = ifelse(server == "FEDERER", as.numeric(receiver.sets), as.numeric(server.sets)))


```


Adding in new variables

```{r newvars}
fed_df <- fed_df %>%
  mutate(speed.diff = NA) %>%
  mutate(oppo.hit.x = NA) %>%
  mutate(oppo.hit.y = NA) %>%
  mutate(oppo.hit.z = NA) %>%
  mutate(oppo.speed = NA)


for (i in 2:nrow(fed_df)) {
  if((fed_df$shot[i] != 1)) { #so that only non-serves are affected
  
  fed_df$speed.diff[i]=fed_df$speed1[i]-fed_df$speed1[i-1]
  #speed difference
  
  fed_df$oppo.hit.x[i]=fed_df$start.x[i-1]
  fed_df$oppo.hit.y[i]=fed_df$start.y[i-1]
  fed_df$oppo.hit.z[i]=fed_df$start.z[i-1]
  fed_df$oppo.speed[i]=fed_df$speed1[i-1]
  #oppo hit
  
  }
}

fed_df <- fed_df %>%
    mutate(sets.diff = fed.sets - opp.sets) %>%
    mutate(games.diff = fed.games - opp.games) %>%
    mutate(points.diff = fed.points - opp.points)
```

Glossary:
lastshot - last shot in the point
fedhit - federer hits the shot
isserver - if server hits the shot
winner - if shot is winner (i.e. point doesn't end in error, as opponent doesn't reach the ball)
retser - if shot is returning serve
speed.diff - difference in current shot speed vs opponents last shot speed (at impact)
speed1 - speed of shot from impact with racquet
oppo.hit - coordinates of opponents previous impact
fed.points/games/sets - fed score up to current point
opp.points/games/sets - opponent score up to current point
points/games/sets.diff - difference in score from Federer's perspective
ballmark.xy - multiplication of ballmark x & y coordinates


```{r noserves}
#make all shots from same side
fed_df <- fed_df %>%
  mutate(start.x = ifelse(start.x <= 0, -start.x, start.x)) %>%
  mutate(oppo.hit.x = ifelse(oppo.hit.x <= 0, -oppo.hit.x, oppo.hit.x)) %>%
  mutate(ballmark.x = ifelse(ballmark.x <= 0, -ballmark.x, ballmark.x))

#remove serves
fed_noser <- fed_df %>% filter(hitpoint != "S") %>%
  dplyr::select(-serveid)

#federer only
fed_only <- fed_noser %>% 
  dplyr::filter(impact.player == "FEDERER")
```


#Visualization - Before We Start Modelling

Here are a series of plots which give us an idea of how each variable affects our response variable - hitting a winner.

###Winners by coordinates of shot, bounce and opponents shot

```{r vis}
library(ggplot2)

#X Co-ordinate of shot
ggplot(fed_only, aes(x=start.x,winner)) + geom_smooth(aes(color = opponent), se=FALSE) + xlab("X Coordinate of Shot") + ylab("Proportion of Winners") + ggtitle("Winners By X Co-Ordinate")

#Y Co-ordinate of shot
ggplot(fed_only, aes(x=start.y,winner)) + geom_smooth(aes(color = opponent), se=FALSE) + xlab("Y Coordinate of Shot") + ylab("Proportion of Winners") + ggtitle("Winners By Y Co-Ordinate")

#Opponents shot before
ggplot(fed_only, aes(x=oppo.hit.x,oppo.hit.y)) + geom_point(aes(color = factor(winner))) + xlab("X Coordinate of Oppo Shot") + ylab("Y Coordinate of Oppo Shot") + ggtitle("Winners By Opponent's Shot Co-Ordinates")

#Y Co-ordinate of Opponent shot
ggplot(fed_only, aes(x=abs(oppo.hit.y),winner)) + geom_smooth(se=TRUE) + xlab("Y Coordinate of Opponent Shot") + ylab("Proportion of Winners") + ggtitle("Winners By Opponent Shot Y Co-Ordinate")

#Shots on x-y plane
ggplot(fed_only, aes(x=start.x,y=start.y)) + geom_point(aes(color = factor(winner))) + xlab("X Coordinate") + ylab("Y Coordinate") + ggtitle("Winners By Shot Co-Ordinates")

#Shots on x-y plane, facet by opponent
ggplot(fed_only, aes(x=start.x,y=start.y)) + geom_point(aes(color = factor(winner))) + facet_wrap(~opponent) + xlab("X Coordinate") + ylab("Y Coordinate") + ggtitle("Winners By Shot Co-Ordinates, Across Opponents")

#Bounce by winner in x-y plane
ggplot(fed_only, aes(x=ballmark.x,y=ballmark.y)) + geom_point(aes(color = factor(winner))) + xlab("X Coordinate") + ylab("Y Coordinate") + ggtitle("Winners By Bounce Co-Ordinates")

#Interaction of x and y bounce to get density
fed_only <- fed_only %>%
mutate(ballmark.xy = ballmark.x * ballmark.y)

ggplot(fed_only) + geom_density(aes(ballmark.xy, group = factor(winner), color=factor(winner)))

#Winners by oppo speed
ggplot(fed_only) + geom_density(aes(oppo.speed, group = factor(winner), color=factor(winner)))
ggplot(fed_only, aes(x=factor(winner),y=oppo.speed,fill=factor(winner))) + geom_boxplot() + ggtitle("Winner by Opponent Speed")

```


###Boxplots for speed and winner across score differences
```{r vis2}
#Box plots across multiple categories
df_long1 <- gather(data = fed_only, key = variables, value = stat, speed1, final.shot, shot)

ggplot(df_long1) + geom_boxplot(aes(x=factor(winner), y=stat, fill=factor(winner))) + facet_wrap(~variables, scale = "free_y") + ggtitle("Differences In Characteristics of Winner/Not")

#Boxplots for speed across games, facet by winner
ggplot(fed_only, aes(x=factor(winner),y=speed1,fill=factor(winner))) + geom_boxplot() + facet_grid(.~ game) + ggtitle("Speed of Shots Across Games")

#Boxplots for speed difference for winner, facet by sets difference
ggplot(fed_only, aes(x=factor(winner),y=speed1,fill=factor(winner))) + geom_boxplot() + facet_grid(.~ sets.diff) + ggtitle("Speed of Shots For Difference in Sets Won")

#Boxplots for speed difference for winner, facet by games difference
ggplot(fed_only, aes(x=factor(winner),y=speed1,fill=factor(winner))) + geom_boxplot() + facet_grid(.~ games.diff) + ggtitle("Speed of Shots For Difference in Games Won")

#Boxplots for speed difference for winner, facet by points difference
ggplot(fed_only, aes(x=factor(winner),y=speed1,fill=factor(winner))) + geom_boxplot() + facet_grid(.~ points.diff) + ggtitle("Speed of Shots For Difference in Points Won")

#Boxplots for speed.diff difference for winner, facet by sets difference
ggplot(fed_only, aes(x=factor(winner),y=speed.diff,fill=factor(winner))) + geom_boxplot() + facet_grid(.~ sets.diff) + ggtitle("Difference in Speed For Difference in Sets Won")

#Boxplots for speed difference for winner, facet by games difference
ggplot(fed_only, aes(x=factor(winner),y=speed.diff,fill=factor(winner))) + geom_boxplot() + facet_grid(.~ games.diff) + ggtitle("Difference in Speed For Difference in Games Won")

#Boxplots for speed difference for winner, facet by points difference
ggplot(fed_only, aes(x=factor(winner),y=speed.diff,fill=factor(winner))) + geom_boxplot() + facet_grid(.~ points.diff) + ggtitle("Difference in Speed For Difference in Points Won")
```

###Winners by point in game

```{r vis3}
#Proportion of winners by forehand/backhand
fed_hand <- na.omit(fed_only) %>%
  group_by(hitpoint,winner) %>%
  summarise(n=n()) %>%
  mutate(proportion = n/sum(n))
ggplot(fed_hand, aes(x=factor(winner),y=proportion,color=factor(winner))) + geom_point() + facet_grid(~hitpoint,labeller = label_both) + ggtitle("Proportion of Winners for Backhand/Forehand Shots")

#Proportion of winners for returning serve
fed_hand <- na.omit(fed_only) %>%
  group_by(retser,winner) %>%
  summarise(n=n()) %>%
  mutate(proportion = n/sum(n))
ggplot(fed_hand, aes(x=factor(winner),y=proportion,color=factor(winner))) + geom_point() + facet_grid(~retser,labeller = label_both) + ggtitle("Proportion of Winners for Returning Serve")


#Winners across points, color by opponent
ggplot(fed_only, aes(x=point,winner)) + geom_smooth(aes(color = opponent), se=FALSE) + xlab("Point") + ylab("Winners as a Proportion of Shots")

#Plotting the winners in each point, across games
ggplot(fed_only, aes(x=point,y=winner)) + geom_smooth(aes(color = factor(game)), se=FALSE) +  xlab("Point") + ylab("Winner") + facet_wrap(~game, scale="free_x") + labs(title="Winners By Point Across Games")

#Plotting the winners in each point, coloured by the set game number, faceted across sets
ggplot(fed_only, aes(x=point,y=winner)) + geom_smooth(aes(color = factor(set)), se=FALSE) +  xlab("Point") + ylab("Winner") + facet_wrap(~game, scale="free_x") + labs(title="Winners By Point Across Sets")
```








# Model 1
### Analysis of Federer's shots only

Response: winner
Covariates: set, game, point, fed.points, opp.points, fed.games, opp.games, fed.sets, opp.sets 

```{r model1}
library(depmixS4)
set.seed(1)
mod.fed_only1 <- depmix(winner ~ 1, transition = ~ set + game + point + fed.points + fed.games + fed.sets, data = fed_only, nstates = 2, family=multinomial("identity"))
fm.fed_only1 <- fit(mod.fed_only1)
summary(fm.fed_only1)
```

State 1 is returning as there is 0 probability of hitting a winner
State 2 is more aggressive, with 0.607 chance of hitting a winner


#Model 2
Additions to Model 1. speed1 as response variable, retser and fhand as covariates

Response: winner, speed1
Covariates: set, game, point, fed.points, opp.points, fed.games, opp.games, fed.sets, opp.sets, retser, fhand, isserver

```{r model2}
library(depmixS4)
mod.fed_only2 <- depmix(list(winner ~ 1, speed1 ~ 1), transition = ~ set + game + point + fed.points + fed.games + fed.sets + retser + fhand + isserver, data = fed_only, nstates = 2, family=list(multinomial("identity"),gaussian()))
fm.fed_only2 <- fit(mod.fed_only2)
summary(fm.fed_only2)
```

State 1 is returning as there is 0.009 chance of hitting a winner, shots are also slower
State 2 is more aggressive, with 0.530 chance of hitting a winner, shots are clearly faster

#Model 3
Scaling speed1 from Model 2

Response: winner, scale(speed1)
Covariates: set, game, point, fed.points, opp.points, fed.games, opp.games, fed.sets, opp.sets, retser, fhand, isserver

```{r model3}
mod.fed_only3 <- depmix(list(winner ~ 1, scale(speed1) ~ 1), transition = ~ set + game + point + fed.points + fed.games + fed.sets + retser + fhand, data = fed_only, nstates = 2, family=list(multinomial("identity"),gaussian()))
fm.fed_only3 <- fit(mod.fed_only3)
summary(fm.fed_only3)
```

Only difference with model 3 from model 2 is that we are scaing the speed of first arc (i.e. speed at impact of ball with racket). This significantly decreases our AIC and makes a much better fit. Probability of hitting a winner from state 2 also increases to 0.584.


### Recall


#Model 4
Adding importance of points score to Model 3

Response: winner, scale(speed1)
Covariates: set, game, point, fed.points, opp.points, fed.games, opp.games, fed.sets, opp.sets, retser, fhand, isserver, importance

```{r model4}
mod.fed_only4 <- depmix(list(winner ~ 1, scale(speed1) ~ 1), transition = ~ set + game + point + fed.points + opp.points + fed.games + opp.games + fed.sets + opp.sets + retser + fhand + importance, data = fed_only, nstates = 2, family=list(multinomial("identity"),gaussian()))
fm.fed_only4 <- fit(mod.fed_only4)
summary(fm.fed_only4)
```

By including the "importance" measure, we increase AIC and also increase probability of hitting a winner from state 1 - which is now 0.008. scale(speed1) parameters are similar to model 3. 

#Model 5
Omitting point, game, set in match from model 4

Response: winner, scale(speed1)
Covariates: fed.points, opp.points, fed.games, opp.games, fed.sets, opp.sets, retser, fhand, isserver, importance


```{r model5}
mod.fed_only5 <- depmix(list(winner ~ 1, scale(speed1) ~ 1), transition = ~ fed.points + opp.points + fed.games + opp.games + fed.sets + opp.sets + retser + fhand + isserver + importance, data = fed_only, nstates = 2, family=list(multinomial("identity"),gaussian()))
fm.fed_only5 <- fit(mod.fed_only5)
summary(fm.fed_only5)
```


Comparison between models:

```{r AICcomp}
AIC(fm.fed_only1)
AIC(fm.fed_only2)
AIC(fm.fed_only3)
AIC(fm.fed_only4)
AIC(fm.fed_only5)

library(lmtest)
lrtest(fm.fed_only2,fm.fed_only1)
lrtest(fm.fed_only3,fm.fed_only2)
lrtest(fm.fed_only4,fm.fed_only3)
lrtest(fm.fed_only5,fm.fed_only4)
```

Model 4 appears to be the best model for this match. This means the time in the match is influential, not just score.

#Predictions with Model 4

```{r predictmodel4}
#Pull out parameters
pars_4 <- getpars(fm.fed_only4)

#Creating temporary transition matrix
#Refresh function: Inputs
#a for time (mxtm), input b for plot number, c for rrep
tr.refresh_4 <- function(a,b,c,d,e,f,g) {
  p11 <- 1/(1+exp(pars_4[[4]]+pars_4[[6]]*a+pars_4[[8]]*b+pars_4[[10]]*c+pars_4[[12]]*d+pars_4[[14]]*e+pars_4[[16]]*f+pars_4[[18]]*g))
  
  p21 <- 1/(1+exp(pars_4[[20]]+pars_4[[22]]*a+pars_4[[24]]*b+pars_4[[26]]*c+pars_4[[28]]*d+pars_4[[30]]*e+pars_4[[32]]*f+pars_4[[34]]*g))
  
  p12 <- 1 - p11
  p22 <- 1 - p21
  
  tr.temp_4 <- c(p11,p21,p12,p22)
  tr.temp_4 <- matrix(tr.temp_4, ncol=2)
}


#Predicted states shell
pr.state_4 <- matrix(0,nrow(fed_only),2)

#State 1
init.state_4 <- matrix(c(pars_4[[1]],pars_4[[2]]),ncol=2) #initial state
a <- fed_only$set[1]
b <- fed_only$game[1]
c <- fed_only$point[1]
d <- fed_only$rscorediff[1]
e <- fed_only$retser[1]
f <- fed_only$oppo.hit.x[1]
g <- fed_only$oppo.hit.y[1]
trans_4 <- tr.refresh_4(a,b,c,d,e,f,g)
pr.state_4[1,] <- init.state_4 %*% trans_4

#States 2 onwards (holding probabilities)
for(i in 2:nrow(fed_only)) {
a <- fed_only$set[i]
b <- fed_only$game[i]
c <- fed_only$point[i]
d <- fed_only$rscorediff[i]
e <- fed_only$retser[i]
f <- fed_only$oppo.hit.x[i]
g <- fed_only$oppo.hit.y[i]
  trans_4 <- tr.refresh_4(a,b,c,d,e,f,g)
  pr.state_4[i,] <- pr.state_4[i-1,] %*% trans_4
}

#Predicted states
predict_df <- data.frame(prob.s1=pr.state_4[,1])

#Making conclusions on state
predict_df <- predict_df %>% rowwise %>%
  mutate(pred.s1 = ifelse(prob.s1 <= 0.5, 0, 1))

test_pred <- cbind(fed_only, predict_df)

#Remember that state 1 is returning, state 2 is winner
```

```{r}
ggplot(test_pred,aes(speed1,winner)) + geom_point(aes(color = pred.s1)) + ggtitle("Model 4 - Response: Winner, Speed")

#Lets think back to the first example speed set. Use that type of visualisation to see how it fits. This here is not so good.
```




```{r visualize winners}
winners <- fed_df %>% filter(lastshot == 1) %>% 
fedwinners <- fed_df %>% filter(impact.player == "FEDERER")

ggplot(fedwinners,aes(ballmark.x,ballmark.y)) + geom_point()
#+ geom_point(aes(color = impact.player))

#+ geom_line(aes(x,y),data=courtlines)
#courtlines <- data.frame(x = c(-11.89, -6.4, 0, 6.4, 11.89), y = c(-5.5, -4.115, 0, 4.115, 5.5))

```

