---
title: "Tennis HMMs"
output: html_notebook
---

#Federer's Matches in Aus Open 2017

```{r loaddata}
setwd("~/Documents/Winter Project/intentions/data")
library(devtools)
install()
load("/Users/Jeremy/Documents/Winter Project/intentions/data/federer2017.RData")

library(ggplot2)
library(tidyr)
library(dplyr)
library(readr)


```

```{r df}
fed_df <- federer2017

fed_df <- arrange(fed_df, matchid, set, game, point, shot)

fed_df$server.score[161:165] = 40
fed_df$receiver.score[161:165] = 0
fed_df$receiver.score[1517] = 0
fed_df$server.score[1517] = 0

#fill in missing scores

fed_df <- fed_df %>%
  mutate(lastshot = ifelse(fed_df$shot == fed_df$final.shot,1,0)) %>%
  mutate(fedhit = ifelse(fed_df$impact.player == "FEDERER",1,0)) %>%
  mutate(isserver = ifelse(fed_df$server == fed_df$impact.player,1,0))

fed_df <- fed_df %>%
  mutate(winner = ifelse(fed_df$ended.in.error == FALSE & fed_df$lastshot == 1,1,0)) %>%
  mutate(retser = ifelse(fed_df$shot == 2,1,0))

for (i in 1:nrow(fed_df)) {
if(fed_df$server.score[i] == "AD") {
  fed_df$server.score[i] = 60
}
  if(fed_df$receiver.score[i] == "AD") {
  fed_df$receiver.score[i] = 60
}
}
#60 means advantage

fed_df$server.score <- as.numeric(fed_df$server.score)
fed_df$receiver.score <- as.numeric(fed_df$receiver.score)

fed_df <- fed_df %>%
  mutate(raw.server.score = NA) %>%
  mutate(raw.receiver.score = NA)

fed_df$raw.receiver.score[1] <- 0
fed_df$raw.server.score[1] <- 0

for (i in 2:nrow(fed_df)) {
  fed_df$raw.server.score[i] = ifelse(fed_df$server.score[i] > fed_df$server.score[i-1], fed_df$raw.server.score [i-1] + 1, ifelse(fed_df$server.score[i] == fed_df$server.score[i-1], fed_df$raw.server.score[i-1], 0))
  
  fed_df$raw.receiver.score[i] = ifelse(fed_df$receiver.score[i] > fed_df$receiver.score[i-1], fed_df$raw.receiver.score [i-1] + 1, ifelse(fed_df$receiver.score[i] == fed_df$receiver.score[i-1], fed_df$raw.receiver.score[i-1], 0))
} 


for (i in 1:ncol(fed_df)) {
  if(is.character(fed_df[,i]) == TRUE) {
    fed_df[,i] <- as.factor(fed_df[,i])
  }
}

```

Adding in new variables

```{r newvars}
fed_df <- fed_df %>%
  mutate(speed.diff = NA) %>%
  mutate(oppo.hit.x = NA) %>%
  mutate(oppo.hit.y = NA) %>%
  mutate(oppo.hit.z = NA)


for (i in 2:nrow(fed_df)) {
  if((fed_df$shot[i] != 1)) { #so that only non-serves are affected
  
  fed_df$speed.diff[i]=fed_df$speed1[i]-fed_df$speed1[i-1]
  #speed difference
  
  fed_df$oppo.hit.x[i]=fed_df$start.x[i-1]
  fed_df$oppo.hit.y[i]=fed_df$start.y[i-1]
  fed_df$oppo.hit.z[i]=fed_df$start.z[i-1]
  #oppo hit
  
  }
}
```

Glossary:
lastshot - last shot in the point
fedhit - federer hits the shot
isserver - if server hits the shot
winner - if shot is winner (i.e. point doesn't end in error, as opponent doesn't reach the ball)
retser - if shot is returning serve
speed.diff - difference in current shot speed vs opponents last shot speed (at impact)
oppo.hit - coordinates of opponents previous impact
rscorediff - raw score difference from Federer's POV

```{r}
#Now filter out serves
fed_noser <- fed_df %>% filter(hitpoint != "S") %>%
  dplyr::select(-serveid)
```

# Model 1
### Begin with the first round match vs Melzer
#### Analysis of Federer's shots only
```{r}
library(dplyr)
fed_mel <- fed_noser %>% 
  filter(matchid == "MS112")  %>%
  mutate(rscorediff = ifelse(server == "FEDERER", server.score - receiver.score, receiver.score - server.score)) %>%
  mutate(fhand = ifelse(hitpoint == "F",1,0)) %>%
  filter(fedhit == "1")
```

We are missing data that tells us about the score up to each shot in the match. This would be nice to give us a better picture of player behavior. Moving on without it.


Response: winner
Covariates: set, game, point, rscorediff

```{r}
library(depmixS4)
set.seed(1)
mod.fed_mel1 <- depmix(winner ~ 1, transition = ~ set + game + point + rscorediff, data = fed_mel, nstates = 2, family=multinomial("identity"))
fm.fed_mel1 <- fit(mod.fed_mel1)
summary(fm.fed_mel1)
```

State 1 is returning as there is 0 probability of hitting a winner
State 2 is more aggressive, with 0.12 chance of hitting a winner

#Model 2
Response: winner, speed1
Covariates: set, game, point, rscorediff, retser, fhand

```{r}
library(depmixS4)
mod.fed_mel2 <- depmix(list(winner ~ 1, speed1 ~ 1), transition = ~ set + game + point + rscorediff + retser + fhand, data = fed_mel, nstates = 2, family=list(multinomial("identity"),gaussian()))
fm.fed_mel2 <- fit(mod.fed_mel2)
summary(fm.fed_mel2)
```

State 1 is returning as there is 0.007 chance of hitting a winner, shots are also slower than 
State 2 is more aggressive, with 0.106 chance of hitting a winner

```{r}
mod.fed_mel3 <- depmix(list(winner ~ 1, scale(speed.diff) ~ 1), transition = ~ set + game + point + rscorediff + retser, data = fed_mel, nstates = 2, family=list(multinomial("identity"),gaussian()))
fm.fed_mel3 <- fit(mod.fed_mel3)
summary(fm.fed_mel3)
```
```{r}
mod.fed_mel4 <- depmix(list(winner ~ 1, speed1 ~ 1), transition = ~ set + game + point + rscorediff + retser + oppo.hit.x + oppo.hit.y, data = fed_mel, nstates = 2, family=list(multinomial("identity"),gaussian()))
fm.fed_mel4 <- fit(mod.fed_mel4)
summary(fm.fed_mel4)
```

Comparison between models:
```{r}
AIC(fm.fed_mel1)
AIC(fm.fed_mel2)
AIC(fm.fed_mel3)
AIC(fm.fed_mel4)
```



```{r visualize winners}
winners <- fed_df %>% filter(lastshot == 1) %>% 
fedwinners <- fed_df %>% filter(impact.player == "FEDERER")

ggplot(fedwinners,aes(ballmark.x,ballmark.y)) + geom_point()
#+ geom_point(aes(color = impact.player))

#+ geom_line(aes(x,y),data=courtlines)
#courtlines <- data.frame(x = c(-11.89, -6.4, 0, 6.4, 11.89), y = c(-5.5, -4.115, 0, 4.115, 5.5))

```

