---
title: "rallies"
author: "Pat Healy"
date: "17 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(devtools)
library(ggplot2)
library(tidyr)
library(dplyr)
library(tidyverse)
```
```{r}
load("data/rallies.RData")
library(readr)
```
Shaping data to recreate the variables used in ("Federer AO 2017.Rmd")'Federer vs Berych 2017' for the AO matches 'Djokovic v Chung 2016', 'Djokovic v Chung 2018', 'Federer v Berdych 2016', 'Federer v Berdych 2018', 'Makarova v Konta 2016','Makarova v Konta 2017','Strycova v Garcia 2016' 'Strycova v Garcia 2017', 'Siegemund v Jankovic 2016','Siegemund v Jankovic 2017', 'Goffin v Thiem 2016''Goffin v Thiem 2017'""

'Siegemund v Jankovic 2017' goes from set 3 to set 6?? 

#### Sorting the matches

```{r}
djo16_df <- filter(rallies, year == 2016, h2h == "DJOKOVIC CHUNG" )
djo16_df <- arrange(djo16_df, matchid, set, game, point, shot)

djo18_df <- filter(rallies, year == 2018, h2h == "DJOKOVIC CHUNG" )
djo18_df <- arrange(djo18_df, matchid, set, game, point, shot)

fed16_df <- filter(rallies, year == 2016, h2h == "FEDERER BERDYCH" )
fed16_df <- arrange(fed16_df, matchid, set, game, point, shot)

fed17_df <- filter(rallies, year == 2017, h2h == "FEDERER BERDYCH" )
fed17_df <- arrange(fed17_df, matchid, set, game, point, shot)

fed18_df <- filter(rallies, year == 2018, h2h == "FEDERER BERDYCH" )
fed18_df <- arrange(fed18_df, matchid, set, game, point, shot)

mak16_df <- filter(rallies, year == 2016, h2h == "MAKAROVA KONTA" )
mak16_df <- arrange(mak16_df, matchid, set, game, point, shot)

mak17_df <- filter(rallies, year == 2017, h2h == "MAKAROVA KONTA" )
mak17_df <- arrange(mak17_df, matchid, set, game, point, shot)

stry16_df <- filter(rallies, year == 2016, h2h == "STRYCOVA GARCIA" )
stry16_df <- arrange(stry16_df, matchid, set, game, point, shot)

stry17_df <- filter(rallies, year == 2017, h2h == "STRYCOVA GARCIA" )
stry17_df <- arrange(stry17_df, matchid, set, game, point, shot)

sie16_df <- filter(rallies, year == 2016, h2h == "SIEGEMUND JANKOVIC" )
sie16_df <- arrange(sie16_df, matchid, set, game, point, shot)

sie17_df <- filter(rallies, year == 2017, h2h == "SIEGEMUND JANKOVIC" )
sie17_df <- arrange(sie17_df, matchid, set, game, point, shot)

thi16_df <- filter(rallies, year == 2016, h2h == "THIEM GOFFIN" )
thi16_df <- arrange(thi16_df, matchid, set, game, point, shot)

thi17_df <- filter(rallies, year == 2017, h2h == "THIEM GOFFIN" )
thi17_df <- arrange(thi17_df, matchid, set, game, point, shot)
```

Starting with fed16 to compare to fed17

Need to compare projected.ballmark.x and start.x stats before reorienting shots. Need to confirm change of terms between data sets first. (no end.x )



#####Recreating variables from Federer AO 2017 for fed16. 
(some columns in rallies data have a different classification than in federer2017 data. Such as 'is.good' replaces 'ended.in.error')(we also now have 'serve' which means we do not need to recreate retser1 or retser2)
```{r}
fed16_df <- fed16_df %>%
  mutate(lastshot = ifelse(shot == final.shot,1,0)) %>%
  mutate(isserver = ifelse(server == impact.player,1,0)) %>%
  mutate(fhand = ifelse(hitpoint == "F",1,0)) %>%
  mutate(opponent = ifelse(server == "FEDERER", receiver, server))

fed16_df <- fed16_df %>%
  mutate(winner = ifelse(is.good == FALSE & lastshot == 1,1,0)) 
  

```

